Initial Analysis

### 2019-11-07 Initial_Analysis.Rmd
Fixed issue with duplicated species status.

#### Library ####
packages <- c(
  "readxl", # Read dataframe
  "data.table", # Read dataframe (Fast!)
  "dplyr", # Data manipulation
  "tidyr", # Data manipulation
  "ggplot2", #Nice grpahs and spatial analysis
  "wesanderson",
  "circlize", # For circular plot
  "ggimage", #for reading images to the circular plot
  # "cowplot",
  # "rgdal",
  # "RColorBrewer",
  # "knitr",
  # "kableExtra",
  "ggrepel", # for nice plot labels
  "ggsflabel", # for nice sf_plots labels
  # "gridExtra",
  # "ggmap",
  # "rgeos",
  # "stringr",
  # "stringer",
  "spdep", # for poly2nb
  "sf", #Spatial analysis 
  "sp", #Spatial analysis 
  "purrr",#Spatial analysis
  "rgdal", #Spatial analysis
  "tools", #Spatial analysis 
  # "png", # For reading plots in chunk codes
  # "grid" # For reading plots in chunk codes
  "parallel", # for parallelization
  "taxize", # For getting species names
  "rfishbase", # for species ecosystem affinity
  "zoo" #for runing mean
)

@@

@@

  No_Status_Data <- Tresholds_Data %>% 
    filter(Met_Criteria != "Yes") %>% 
    select(TaxonKey,Territory) %>% 
    mutate(Status = "No Status")

@@

  # Get Variables to perform status metric
  Status_Data <- Data %>% 
    anti_join(No_Status_Data, # Removes no status data
              by = c("TaxonKey","Territory")
    ) %>% 
    # Get Max catch and max catch year
    group_by(TaxonKey,Territory) %>% 
    top_n(1,RMean) %>% 
    rename(Max_Peak = RMean,
           Y_Max_Peak = Year) %>% 
    # Get Min catch after peak catch
    left_join(Data,
              by = c("TaxonKey","Territory")
    ) %>% 
    filter(Year > Y_Max_Peak) %>% # Filter years after peak year
    group_by(TaxonKey,Territory) %>% 
    top_n(-1,RMean) %>% # find min value
    rename(Post_Max_Min = RMean,
           Y_Post_Max_Min = Year) %>% 
    left_join(Data,
              by = c("TaxonKey","Territory")
    ) %>% 
    # Final Species Status Evaluation 
    mutate(Status = ifelse(Year > Y_Post_Max_Min & Post_Max_Min < (Max_Peak*0.10) & ((RMean > (Max_Peak*0.10) & RMean < (Max_Peak*0.50))),"Rebuilding",
                           ifelse((Year < Y_Max_Peak & RMean <= (Max_Peak*0.50) | Y_Max_Peak == max(Year)),"Developing",
                                  ifelse(RMean > (Max_Peak*0.50),"Max Exploited",
                                         ifelse(Year > Y_Max_Peak & ((RMean > (Max_Peak*0.10) & RMean < (Max_Peak*0.50))),"Over Exploited",
                                                ifelse(Year > Y_Max_Peak & RMean < (Max_Peak*0.10),"Collapsed",
                                                       "No Status")
                                         )
                                  )
                           )
    )) %>% 
    filter(Year > 2004) %>% # last 10 years of data
    group_by(TaxonKey,Territory,Status) %>% 
    summarise(n_Status= n()) %>% 
    group_by(TaxonKey,Territory) %>% 
    top_n(1,n_Status) %>%
    select(-n_Status)

@@

  # If both datasets have value RETURN BOTH
  if(nrow(No_Status_Data) > 0 & nrow(Status_Data) > 0){

    Fun_Result <- bind_rows(No_Status_Data,
                            Status_Data)
  }

  # If only No status data, then just return No Status Data
  if(nrow(No_Status_Data) > 0 & nrow(Status_Data) == 0){
    return(No_Status_Data)
  }else{
    return(Status_Data)
  }
}

@@

# Save data
write.csv(Species_Status,
          paste(Results_Path,"Species_Status.csv",sep=""),
                row.names = F)

@@

# Read Species Status Data (Sub-routine Species Status)
Status_Data <- fread(paste(Data_Path,"Distribution_Data/Species_Status.csv", sep =""))
# Status_Data <- Species_Status

@@

Status_Data_Plot <- Tresholded_T_Data %>% 
  # mutate(TaxonKey = as.character(TaxonKey)) %>%
  # rename(Territory = CT_fishing_entity) %>% 
  left_join(Status_Data, 
            by =c("Territory","TaxonKey")
  ) %>%
  mutate(Status = ifelse(is.na(Status),"No Status",Status)) %>% 
  group_by(Territory,Status) %>% 
  summarise(value = length(unique(TaxonKey))) %>% 
  left_join(UN_Regions, by ="Territory") %>%
  mutate(Georegion = ifelse(is.na(georegion),"Other",paste(georegion))) %>% 
  rename(FishingRegion=Territory) %>% 
  left_join(Clean_sau_eez_fes) %>% 
  group_by(individual = fishing_entity,
           group = Georegion,
           observation = Status) %>% 
  summarise(Total_trans = sum(value,na.rm=T)) %>% 
  mutate(value = ifelse(Total_trans > 100,100,Total_trans)) %>% 
  filter(!is.na(individual)) %>% 
  select(individual,observation,group,value) %>% 
  arrange(group,value) %>% 
  mutate(value = ifelse(observation %in% c("Collapsed","Over Exploited"),value*-1,value))

@@

>>>>

#### Library ####
packages <- c(
  "readxl", # Read dataframe
  "data.table", # Read dataframe (Fast!)
  "dplyr", # Data manipulation
  "tidyr", # Data manipulation
  "ggplot2", #Nice grpahs and spatial analysis
  "wesanderson",
  "tidyverse",
  "circlize", # For circular plot
  "ggimage", #for reading images to the circular plot
  # "cowplot",
  # "rgdal",
  # "RColorBrewer",
  # "knitr",
  # "kableExtra",
  "ggrepel", # for nice plot labels
  "ggsflabel", # for nice sf_plots labels
  # "gridExtra",
  # "ggmap",
  # "rgeos",
  # "stringr",
  # "stringer",
  "spdep", # for poly2nb
  "sf", #Spatial analysis 
  "sp", #Spatial analysis 
  "purrr",#Spatial analysis
  "rgdal", #Spatial analysis
  "tools", #Spatial analysis 
  # "png", # For reading plots in chunk codes
  # "grid" # For reading plots in chunk codes
  "parallel", # for parallelization
  "taxize", # For getting species names
  "rfishbase", # for species ecosystem affinity
  "zoo" #for runing mean
)

@@

Spp <- 600006
Coord <- CoorG

@@

  No_Status_Data <- Tresholds_Data %>% 
    filter(Met_Criteria != "Yes") %>% 
    select(TaxonKey,Territory) %>% 
    mutate(Status = "No Trend")

@@

  # Get Variables to perform status metric
  Status_Data <- Data %>% 
    anti_join(No_Status_Data, # Removes no status data
              by = c("TaxonKey","Territory")
    ) %>% 
    # Get Max catch and max catch year
    group_by(TaxonKey,Territory) %>% 
    top_n(1,RMean) %>% 
    rename(Max_Peak = RMean,
           Y_Max_Peak = Year) %>% 
    # Get Min catch after peak catch
    left_join(Data,
              by = c("TaxonKey","Territory")
    ) %>% 
    filter(Year > Y_Max_Peak) %>% # Filter years after peak year
    group_by(TaxonKey,Territory) %>% 
    top_n(-1,RMean) %>% # find min value
    group_by(TaxonKey,Y_Max_Peak,Territory,Max_Peak,RMean) %>%  # Solves for years with the same catch data (e.g. Cayman Isl 2013-2014)
    summarise(Year = min(Year)) %>% 
    rename(Post_Max_Min = RMean,
           Y_Post_Max_Min = Year) %>% 
    left_join(Data,
              by = c("TaxonKey","Territory")
    ) %>% 
    # Final Species Status Evaluation 
    mutate(Status = ifelse(Year > Y_Post_Max_Min & Post_Max_Min < (Max_Peak*0.10) & ((RMean > (Max_Peak*0.10) & RMean < (Max_Peak*0.50))),"Rebuilding",
                           ifelse((Year < Y_Max_Peak & RMean <= (Max_Peak*0.50) | Y_Max_Peak == max(Year)),"Developing",
                                  ifelse(RMean > (Max_Peak*0.50),"Max Exploited",
                                         ifelse(Year > Y_Max_Peak & ((RMean > (Max_Peak*0.10) & RMean < (Max_Peak*0.50))),"Over Exploited",
                                                ifelse(Year > Y_Max_Peak & RMean < (Max_Peak*0.10),"Collapsed",
                                                       "No Trend")
                                         )
                                  )
                           )
    )
    )

  # Determine the final status using the top category over the last 10 years
  Determine_Status_Data <- Status_Data %>% 
    filter(Year > 2004) %>% # last 10 years of data (2005-2014)
    group_by(TaxonKey,Territory,Status) %>%
    summarise(n_Status= n()) %>%
    group_by(TaxonKey,Territory) %>%
    top_n(1,n_Status)

  # Fix cases where you have equal categories.
  Duplicates <- Determine_Status_Data %>% 
    ungroup() %>% 
    janitor::get_dupes(Territory) %>% 
    select(TaxonKey,Territory) %>% 
    pull(Territory) %>% 
    unique()

  # In this case we sue the last year's category
  Duplicates_Status <- Status_Data %>% 
  filter(Territory %in% Duplicates,
         Year == 2014) %>%
    ungroup() %>%
    select(TaxonKey,Territory,Status)

  # Build Final Status database
  Final_Status_Data <- Determine_Status_Data %>% 
    filter(!Territory %in% Duplicates) %>% 
    bind_rows(Duplicates_Status)

@@

  # If both datasets have value RETURN BOTH
  if(nrow(No_Status_Data) > 0 & nrow(Final_Status_Data) > 0){

    Fun_Result <- bind_rows(No_Status_Data,
                            Final_Status_Data)
  }

  # If only No status data, then just return No Status Data
  if(nrow(No_Status_Data) > 0 & nrow(Final_Status_Data) == 0){
    return(No_Status_Data)
  }else{
    return(Final_Status_Data)
  }
}

@@

# Save data
# readr::write_csv(Species_Status,
#           "Species_Status.csv")
# head(Species_Status)
```

@@

# Read Species Status Data (Sub-routine Species Status)
Status_Data <- fread(paste(Results_Path,"Species_Status.csv", sep =""))
# Status_Data <- Species_Status

@@

Status_Data_Plot <- Tresholded_T_Data %>% 
  mutate(TaxonKey = as.character(TaxonKey)) %>%
  # rename(Territory = CT_fishing_entity) %>% 
  left_join(Species_Status, 
            by =c("Territory","TaxonKey")
  ) %>%
  mutate(Status = ifelse(is.na(Status),"No Status",Status)) %>% 
  group_by(Territory,Status) %>% 
  summarise(value = length(unique(TaxonKey))) %>% 
  left_join(UN_Regions, by ="Territory") %>%
  mutate(Georegion = ifelse(is.na(georegion),"Other",paste(georegion))) %>% 
  rename(FishingRegion=Territory) %>% 
  left_join(Clean_sau_eez_fes) %>% 
  group_by(individual = fishing_entity,
           group = Georegion,
           observation = Status) %>% 
  summarise(Total_trans = sum(value,na.rm=T)) %>% 
  mutate(value = ifelse(Total_trans > 100,100,Total_trans)) %>% 
  filter(!is.na(individual)) %>% 
  select(individual,observation,group,value) %>% 
  arrange(group,value) %>% 
  mutate(value = ifelse(observation %in% c("Collapsed","Over Exploited"),value*-1,value))

@@

### Number of Collapsed USA species

```{r Collapsed_USA, eval = F ,echo = T,warning = F,message = F}

#  At the end, the ram legacy and the SAU are tide in results...

USA_Collapsed <- Species_Status %>% 
  filter(str_detect(Territory, pattern="USA")#,
         # Status == "Collapsed"
  ) %>% 
  ungroup() %>% 
  mutate(TaxonKey = as.numeric(TaxonKey)) %>%
  left_join(Exploited_Species)

# Comparring to RAM legacy in the W pacific coast

# Sabelfish match
# Pacifc sardine no data ram
# Chillipepper match
# Lincod no match althou circular path in RAM
# Dover no match by far...
# Chub mackerel... match
# Petral metch
# Black Rockfish No mach...ish
# Widow rockfish no match
# Starry flounder no match
# Rex sole no data RAM
# Pacific ocean percCollapsed Vs Overfished
#
